---
title: "Biodiversity Metrics in Ag Landscapes"
format: gfm
editor: visual
editor_options: 
  chunk_output_type: console
execute: 
  message: false
  warning: false
---

## Import National Landcover Data (NLCD)

```{r}
library(tidyverse)
library(landscapemetrics)     # landscape metrics calculation
library(raster)               # spatial raster data reading and handling
library(gt)

wash <- raster("data/landcover_wa.tiff") 
iowa <- raster("data/landcover_ia.tiff") 
georgia <- raster("data/landcover_ga.tiff") 
kansas <- raster("data/landcover_ks.tiff")
```

Let's look at the four farms in this example.

```{r}
#| fig-height: 8
par(mfrow = c(2,2))
plot(wash)
title("WA")
plot(iowa)
title("IA")
plot(kansas)
title("KS")
plot(georgia)
title("GA")
```

## Shannon's diversity index

It is a widely used metric in biodiversity and ecology and takes both the number of classes and the abundance of each class into account.

$$SHDI = − \sum_{i = 1}^m ​P_i​ \times\ \ln(P_i​)$$

where $P_i$ is the proportion of class $i$, and $m$ is the number of classes.

```{r}
#| tbl-cap: Shannon's diversity index
map(c(wash, kansas, georgia, iowa), lsm_l_shdi) |> 
  list_rbind() |> 
  mutate(farm = c("wash", "kansas", "georgia", "iowa")) |> 
  gt()
```

## Shannon's evenness index

SHEI is a 'Diversity metric'. It is the ratio between the actual Shannon's diversity index and and the theoretical maximum of the Shannon diversity index. It can be understood as a measure of dominance.

$$SHEI = \frac{− \sum \limits_{i = 1}^m ​P_i​ \times\ \ln(P_i​)}{\ln(m)}$$

where $P_i$ is the proportion of class $i$, and $m$ is the number of classes.

```{r}
#| tbl-cap: Shannon's evenness index
map(c(wash, kansas, georgia, iowa), lsm_l_shei) |> 
  list_rbind() |> 
  mutate(farm = c("wash", "kansas", "georgia", "iowa")) |> 
  gt()
```

## Simpson's diversity index

SIDI is a 'Diversity metric'. It is widely used in biodiversity and ecology. It is less sensitive to rare class types than Shannon's diversity index. It can be interpreted as the probability that two randomly selected cells belong to the same class.

$$SIDI = 1 − \sum_{i = 1}^m ​P_i^2$$

```{r}
#| tbl-cap: Simpson's diversity index
map(c(wash, kansas, georgia, iowa), lsm_l_sidi) |> 
  list_rbind() |> 
  mutate(farm = c("wash", "kansas", "georgia", "iowa")) |> 
  gt()
```

### References

Hesselbarth, M.H.K., Sciaini, M., With, K.A., Wiegand, K., Nowosad, J. 2019. landscapemetrics: an open-source R tool to calculate landscape metrics. Ecography 42:1648-1657(ver. 0).

McGarigal, K., SA Cushman, and E Ene. 2012. FRAGSTATS v4: Spatial Pattern Analysis Program for Categorical and Continuous Maps. Computer software program produced by the authors at the University of Massachusetts, Amherst. Available at the following web site: https://www.umass.edu/landeco/

Shannon, C., and W. Weaver. 1949. The mathematical theory of communication. Univ. IllinoisPress, Urbana

Simpson, E. H. 1949. Measurement of diversity. Nature 163:688

## Other things to explore

-   [this link](https://smalltownbigdata.github.io/feb2021-landcover/feb2021-landcover.html)
-   [this link](https://gatesdupontvignettes.com/2019/06/16/nlcd-velox-dplyr.html)
-   and [this link](https://jakubnowosad.com/posts/2022-02-17-lsm-bp3/#reading-the-data)
