---
title: "Biodiversity Metrics in Ag Landscapes"
format: gfm
output-file: README.md
editor: visual
editor_options: 
  chunk_output_type: console
execute: 
  message: false
  warning: false
---

The Semi-Natural Habitat (SNH) composite score is

$$SNH = (CP + Veg) \times{(1 + SizeRatio) + Configuration + Management}$$

where\
$CP$ = Conservation practice base score,\
$Veg$ = Vegetative diversity score

therefore, the proposed biodiversity metric would be

$$
New\ Metric = (SNH + CoverCrops + Tillage) \times{LC} + PM + Rotation - Conversion
$$

where\
LC = Landscape Complexity score ([Examples of landscape complexity calculations])\
PM = Pest Management score

To explore the point values and the calculations under different farming scenarios, please explore these spreadsheets and the accompanying sensitivity analysis:

-   [Scenario 1](scenario/scenario-01.xlsx)

-   [Scenario 2](scenario/scenario-02.xlsx)

-   [Scenario 3](scenario/scenario-03.xlsx)

## Comments from reviewers

> Overall I must applaud your team on the shift in focus from the previous document. It is very impressive and I believe a much more robust framework. However it does open up some doors that are complicated to walk through and some that are very difficult to close.

> Could we call it Biodiversity Potential Index or Conservation Potential Index?
>
> I would advise moving away from the term Habitat. Your focus is on Biodiversity and you do not need to invoke 'habitat' to get there.\
> \
> Technically there is no such thing as 'wildlife habitat'. There is quail habitat, rabbit habitat, or deer habitat.

> I think your metric should be driven by semi-natural vegetation.
>
> The literature is abundantly clear that the amount of non-crop vegetation is what is driving biodiversity in ag landscapes.

> It is only valuable to differentiate between cover types if you are going to assign different values to different cover types. That is a noble endeavor but will present some challenges because different cover types may be more important in different parts of the country.

> Be careful with Rangeland Analysis Platform as it is not as good east of the Mississippi river as it is west of it.

> "edge" is double-edged sword in wildlife biodiversity.

> if grower is operating an hot bed of biodiversity already, then the conservation-friendly actions they take will have a larger impact, depending on what metric is in question. However, if a VERY conservation-friendly farmers is operating in an extremely bio-hostile landsdape, then their actions will have considerably less impact on the landscape. ... In other words the landscape context DOES matter, and I support quantifying it, BUT how that information is used \[in the calculation\] is unclear.

## Concerns and questions for reviewers

-   This metrics revision still does not capture the value of the habitat creation from a flooded rice field
-   What can be categorized as a cover crop?
    -   Could we consider using months of a ***living/rooted*** land cover and months/quality of a ***residue*** cover
-   Does this revision represent an overall improvement according to the available science
-   Should we or how would we track temporal changes in the diversity index?
    -   Farm-by-farm analysis could be more data heavy, so another option is county based reporting with FSA
-   What should the point values be in each section of the calculation, for different practices?
-   How should we present a biodiversity result?
    -   Quantitative (e.g., 0.76, 1.21)

    -   Qualitative (e.g., low, moderate)

## Questions to be weighed by committee

### **Scope questions**

(relating to major departures from the previous metric)

-   Does the Committee agree on the need to emphasize the value of non-cropland, semi-natural habitat (SNH)?

    -   edge of field practices, adjacent uncropped land, in-field (ephemeral) wetland habitat, etc.

-   Does the Committee want to include landscape complexity as a weighting factor for habitat-provisioning practices (SNH, cover crops, residue, etc.)?

    -   Landscape complexity is largely out of the growers' operational control (like soil types in the erosion and GHG metrics), *but* may introduce more accuracy into the metric.

-   Does the Committee agree on the need to provide greater weight for pest management and increase the granularity of questions?

-   How should the metric results be presented to the use?

    -   0-100%

    -   0 to number less than 100%

    -   poor/moderate/high quality

    -   etc.?

### **Implementation questions**

(relating to point values and question design)

-   Landscape complexity

    -   Shannon vs. Simpson diversity index?

    -   What is landscape-scale?

        -   1km around the field, 5km around the field?

        -   From the center of the field or from field borders?

-   Cover crops

    -   Are the point values correct?

        -   Is there sufficient research to support these values?

    -   Do winter wheat and alfalfa receive the same point values?

-   Point values of:

    -   SNH base practices

    -   Vegetative diversity

        -   Can users reliably differentiate between cover types?

    -   Management

        -   Does asking about management of semi-natural vegetation lead to double counting in the vegetative diversity question, since the purpose of the management is often to maintain stand diversity and vigor?

        -   What is the appropriate timeline for considering management (in current year, in past three years, in past five years)?

-   Pest management

    -   Are these the right BMPs and point values?

    -   How does a user respond to this question if they perform multiple crop protectant applications, i.e. proactive scouting for an mid-season application, but not an early-season application?

-   Cropland conversion

    -   Can NLCD reliably detect land use change? Would other land cover maps be better fit for purpose?

-   Crop rotation

    -   How should the crop rotation question look at alfalfa (or other perennials available in the rotation builder)? What about (chem) fallow years?

------------------------------------------------------------------------

## Examples of landscape complexity calculations

We could consider using a landscape complexity score as a sort of multiplier in the overall equation. Here are some examples:

### Import National Landcover Data (NLCD)

```{r}
library(tidyverse)
library(landscapemetrics) # landscape metrics calculation
library(raster)           # spatial raster data reading and handling
library(knitr)            # markdown tables

wash <- raster("data/landcover_wa.tiff") 
iowa <- raster("data/landcover_ia.tiff") 
georgia <- raster("data/landcover_ga.tiff") 
kansas <- raster("data/landcover_ks.tiff")
```

Let's look at the four farms in this example.

```{r}
#| fig-height: 8
#| code-fold: true
par(mfrow = c(2,2))
plot(wash)
title("WA")
plot(iowa)
title("IA")
plot(kansas)
title("KS")
plot(georgia)
title("GA")
```

## Shannon's diversity index

It is a widely used metric in biodiversity and ecology and takes both the number of classes and the abundance of each class into account.

$$SHDI = − \sum_{i = 1}^m ​P_i​ \times\ \ln(P_i​)$$

where $P_i$ is the proportion of class $i$, and $m$ is the number of classes.

```{r}
#| tbl-cap: Table 1 Shannon's diversity index
map(c(wash, kansas, georgia, iowa), lsm_l_shdi) |> 
  list_rbind() |> 
  mutate(farm = c("wash", "kansas", "georgia", "iowa")) |> 
  kable(digits = 2)
```

## Shannon's evenness index

SHEI is a 'Diversity metric'. It is the ratio between the actual Shannon's diversity index and and the theoretical maximum of the Shannon diversity index. It can be understood as a measure of dominance.

$$SHEI = \frac{− \sum \limits_{i = 1}^m ​P_i​ \times\ \ln(P_i​)}{\ln(m)}$$

where $P_i$ is the proportion of class $i$, and $m$ is the number of classes.

```{r}
#| tbl-cap: Table 2 Shannon's evenness index
map(c(wash, kansas, georgia, iowa), lsm_l_shei) |> 
  list_rbind() |> 
  mutate(farm = c("wash", "kansas", "georgia", "iowa")) |> 
  kable(digits = 2)
```

## Simpson's diversity index

SIDI is a 'Diversity metric'. It is widely used in biodiversity and ecology. It is less sensitive to rare class types than Shannon's diversity index. It can be interpreted as the probability that two randomly selected cells belong to the same class.

$$SIDI = 1 − \sum_{i = 1}^m ​P_i^2$$

```{r}
#| tbl-cap: Table 3 Simpson's diversity index
map(c(wash, kansas, georgia, iowa), lsm_l_sidi) |> 
  list_rbind() |> 
  mutate(farm = c("wash", "kansas", "georgia", "iowa")) |> 
  kable(digits = 2)
```

### References

Hesselbarth, M.H.K., Sciaini, M., With, K.A., Wiegand, K., Nowosad, J. 2019. landscapemetrics: an open-source R tool to calculate landscape metrics. Ecography 42:1648-1657(ver. 0).

McGarigal, K., SA Cushman, and E Ene. 2012. FRAGSTATS v4: Spatial Pattern Analysis Program for Categorical and Continuous Maps. Computer software program produced by the authors at the University of Massachusetts, Amherst. Available at the following web site: https://www.umass.edu/landeco/

Shannon, C., and W. Weaver. 1949. The mathematical theory of communication. Univ. IllinoisPress, Urbana

Simpson, E. H. 1949. Measurement of diversity. Nature 163:688

## Other things to explore

-   [this link](https://smalltownbigdata.github.io/feb2021-landcover/feb2021-landcover.html)
-   [this link](https://gatesdupontvignettes.com/2019/06/16/nlcd-velox-dplyr.html)
-   and [this link](https://jakubnowosad.com/posts/2022-02-17-lsm-bp3/#reading-the-data)
